
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
# load DESeq2 library (this takes some time)
#if (!requireNamespace("BiocManager", quietly = TRUE))
    #install.packages("BiocManager")
#BiocManager::install("DESeq2")


```{r} 
count_data <- read.csv("count_data.csv", row.names = 1)
sample_info <- read.csv("sample_info.csv", row.names = 1)
dds <- DESeqDataSetFromMatrix(countData = count_data,
                              colData = sample_info,
                              design = ~ tissueType + individualID)
```{r}
dds <- dds[rowSums(counts(dds)) >= 10, ]
num_genes <- nrow(dds)
num_genes
dds <- DESeq(dds)



Sort genes in each vector by adjusted p-values in an ascending order.

```{r}
result <- results(dds, contrast = c("tissueType", "primary_tumor", "normal"))
upregulated_genes <- subset(result, log2FoldChange > 2 & padj < 0.05)
upregulated_genes <- upregulated_genes[order(upregulated_genes$padj), ]
upregulated_genes <- rownames(upregulated_genes)
downregulated_genes <- subset(result, log2FoldChange < -2 & padj < 0.05)
downregulated_genes <- downregulated_genes[order(downregulated_genes$padj), ]
downregulated_genes <- rownames(downregulated_genes)
upregulated <- length(upregulated_genes)
downregulated <- length(downregulated_genes)
upregulated_genes
downregulated_genes
upregulated
downregulated


###  Plot normalized counts for the most significant gene

find out the most significant gene with the smallest adjusted p-value. plot the normalized counts for this gene in 3 different tissue types.

```{r}
res <- result[order(result$padj), ]
most_sig_gene <- rownames(res)[1]
most_sig_gene
log2FoldChange <- res$log2FoldChange[1]
regulation <- ifelse(log2FoldChange > 0, "up-regulated", "down-regulated")
regulation
plotCounts(dds, gene = most_sig_gene, intgroup = "tissueType")

### Volcano plot


```{r}
mycols <- ifelse(result$log2FoldChange > 2 & result$padj < 0.05, "red",
                 ifelse(result$log2FoldChange < -2 & result$padj < 0.05, "blue", "black"))
plot(result$log2FoldChange, result$padj, col = mycols, pch = 20,
     xlab = "Log2FoldChange", ylab = "Adjusted p-value (padj)", main = "Volcano Plot")
abline(v = c(-2, 2), col = "gray", lty = 2)
abline(h = 0.05, col = "gray", lty = 2)
```

### EdgeR analysis


```{r, include=FALSE}
# Load libraries (please install them if you haven't installed them)
#if (!requireNamespace("BiocManager", quietly = TRUE))
    #install.packages("BiocManager")
BiocManager::install("limma")
BiocManager::install("edgeR")

library(limma)
library(edgeR)
```

```{r}
group <- factor(sample_info$tissueType)
dge <- DGEList(counts = count_data, group = group)
keep <- filterByExpr(dge, group = group)
dge <- dge[keep, , keep.lib.sizes = FALSE]
dge <- calcNormFactors(dge)
dge <- estimateDisp(dge)
test <- exactTest(dge, pair = c("normal", "primary_tumor"))
top_genes <- topTags(test, n = 10)
top_genes$table
```
